---
title: "probability"
author: "Alexa Kelly"
date: "6/19/2020"
output: html_document
---


Harvard Data Sceince: Probability

One ball will be drawn at random from a box containing: 3 cyan balls, 5 magenta balls, and 7 yellow balls.

Possibility of Cyan
```{r }
cyan <- 3
magenta <- 5
yellow<-7

#probability of the ball being cyan: (with replacement)
p_1 = cyan/(cyan + magenta + yellow)
p_1
```

Probability of the ball being not cyan: (with replacement)
```{r }
p_2 = (magenta + yellow)/(cyan + magenta + yellow)
p_2

```

Probability (without replacement): first draw is cyan and that the second draw is not cyan?
A : the first draw is cyan
B : the second draw is not cyan

Using math equations, we have that P(A) = 3/15 = 1/5 

```{r }
#probability of not choosing a cyan ball on the second draw without replacement.
p_3 <- 1 - (cyan - 1) / (cyan + magenta + yellow - 1) # P(B/A) = 12/14 = 6 / 7 so ... 
#P(A and B ) = P(A) * P(B/A) = 1/5 * 6/7 = 6/35

#so the probability that the first draw is cyan and the second draw is not cyan:
p_1 * p_3

```

Sampling with replacement.
now, after taking the first draw and recording the color, return it back to the box and shake the box. 
```{r }
#probability the first draw is cyan and that the second draw is not cyan:
#P(A) = 1/5       # we called this p_1
p_4 <- 1- cyan/ (cyan + magenta + yellow)     # P(C) = 1- 1/5 = 4/5
p_1 * p_4    # P(A and C) = P(A) * P(B) = 1/5 * 4/5 = 4/25
```

