---
title: "discrete"
author: "Alexa Kelly"
date: "6/24/2020"
output: html_document
---

Harvard Data Sceince: Probability


Possibility of Cyan

One ball will be drawn at random from a box containing: 3 cyan balls, 5 magenta balls, and 7 yellow balls.
```{r setup }
cyan <- 3
magenta <- 5
yellow<-7
```


Probability of the ball being cyan (p_1) verses the probability of the ball being not cyan (p_2), WITH replacement.
```{r }
p_1 = cyan / (cyan + magenta + yellow)
p_1

p_2 = (magenta + yellow) / (cyan + magenta + yellow)
p_2
```


Probability, without replacement: first draw is cyan and that the second draw is not cyan?
A : the first draw is cyan
B : the second draw is not cyan

Using using math equations, using the dependent probability equation, we have that:
P(A) = 3/15 = 1/5 

To find the probability of not choosing a cyan ball on the second draw without replacement, we need P(A⋂B) =  P(A) * P(B/A) = 1/5 * 6/7 = 6/35

where P(B/A) = 12/14 = 6 / 7 so ... 
P(A) is p_1, the first probability we calculated. We can define p_3 as P(B/A).


```{r }
#probability of not choosing a cyan ball on the second draw without replacement.
p_3 <- 1 - (cyan - 1) / (cyan + magenta + yellow - 1) 

#so the probability that the first draw is cyan and the second draw is not cyan:
p_1 * p_3
```


Now let's sample with replacement. After taking the first draw and recording the color, return it back to the box and shake the box. 

The Multiplication Rule for independent events:
P(A⋂B⋂C)= P(A and B and C) = P(A) * P(B) * P(C)

The probability the first draw is cyan and that the second draw is not cyan is P(A) = 1/5. As mentioned earlier, we called this p_1.

Now we define P(C), named p_4 here. 
P(C) = 1- 1/5 = 4/5.

So, to sample with replacement, we have:  P(A and C) = P(A) * P(C) = 1/5 * 4/5 = 4/25

```{r }
p_4 <- 1- cyan/ (cyan + magenta + yellow) 
p_1 * p_4 
```
