---
title: "monte carlo"
author: "Alexa Kelly"
date: "6/19/2020"
output: html_document
---


Harvard Data Science: Probability Course 
 "Monte Carlo simulations"
 
Create an urn with 2 red and 3 blue beads
```{r setup, include=FALSE}
#use rep function to generate the urn, 
beads <- rep(c("red", "blue"), times = c(2,3))    

beads    # view beads object
sample(beads, 1)    # sample 1 bead at random

B <- 10000    # number of times to draw 1 bead
events <- replicate(B, sample(beads, 1))    # draw 1 bead, B times
tab <- table(events)    # make a table of outcome counts
tab    # view count table
prop.table(tab)    # view table of outcome proportions
```

A more accurate use of stats for replacing the beads would be:
```{r }
events2 <- sample(beads, B, replace= TRUE)
prop.table(table(events2))

#  the vector beads from ^:
beads <- rep(c("red", "blue"), times = c(2,3))
beads

```

To find the probability of drawing a blue bead at random, take the mean:
```{r }
mean(beads == "blue")
```
Clearly there is a .6 chance of picking a blue bead as the first.

Finding the outcomes of the other 4 is as follows:
```{r }
x <- sample(beads, 5)
x[2:5]                
```
These events are not independent! each pick affects the probability of the next color.
